---
title: "Food Consulting"
author: "Cooper Riggs & Stockton Nelson"
format: pdf
editor: visual
---

## Setup

```{r}
#Downloading Libraries
library(tidyverse)

#Bringing in Data
food <- vroom::vroom("FoodExpenses.txt") |>
  janitor::clean_names() |>
  rename(eating = eating_out)
```

## Problem 2 (Cooper)

### Relationship Between Income and Food Expenditures

```{r}

```

### Healthy Economy

```{r}

```

### Predictions

```{r}

```

## Problem 3 (Stockton)

### Problems with Standard Regression

```{r}
#Making Linear Model
eqlm <- lm(eating ~ ., data = food)

#Making Scatterplot of Birth Weight and Mother's Age with a Prediction Interval
preds <- predict(eqlm, interval = "prediction")
plot_df <- food |>
  mutate(
    fit = preds[,1],
    lwr = preds[,2],
    upr = preds[,3]
  )
ggplot(data = plot_df, mapping = aes(x = income, y = eating)) +
  geom_point() +
  geom_line(aes(y = fit), color = "blue", size = 1.125) +
  geom_line(aes(y = lwr), color = "red", size = 1.125) + 
  geom_line(aes(y = upr), color = "red", size = 1.125) +
  labs(x = "Income in Thousands ($)",
       y = "Eating Out ($)",
       title = "Scatterplot",
       subtitle = "Income and Money Spent Eating Out")

#Saving the Scatterplot
ggsave("Scatterplot_SLR.png", width = 8, height = 6, dpi = 300)
```

```{r}
#Checking Equal Variance Assumption
stresids <- MASS::stdres(eqlm)
ggplot(data = data.frame(fit = fitted(eqlm), res = stresids),
       aes(x = fit, y = res)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "blue", size = 1.125) +
  labs(x = "Fitted",
       y = "Residuals")

#Saving the Fitted Vs Residuals Plot
ggsave("Fitted_VS_Residuals_SLR.png", width = 8, height = 6, dpi = 300)
```

```{r}
#Running BP Test
resid_df <- food |>
  mutate(r = stresids) |>
  select(-eating)
lmtest::bptest(formula = r ~ ., data = resid_df)
```

### Validate Unequal Variance Model

#### Linearity

```{r}
#Making Scatterplot of Birth Weight and Mother's Age
ggplot(data = plot_df, mapping = aes(x = income, y = eating)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) + 
  labs(x = "Income in Thousands ($)",
       y = "Eating Out ($)",
       title = "Scatterplot",
       subtitle = "Income and Money Spent Eating Out")

#Saving the Scatterplot
ggsave("Scatterplot_UV.png", width = 8, height = 6, dpi = 300)
```

#### Independence

With independence, we have to look at the method by which the data we used was collected. Given what we know about the dataset, there is no reason to believe that the amount spent on fast food by one household would affect the amount spent on fast food by another household. Given what we know, we are able to conclude independence.

#### Normality

```{r}
#Fitting a Heteroskedastic Model
uvlm <- gamlss::gamlss(eating ~ ., sigma.formula = ~income, data = food)

#Getting the Standardized Residuals
stresids <- residuals(uvlm, what = "z-scores")

#Making Histogram
ggplot() +
  geom_histogram(aes(x = stresids, y = after_stat(density))) +
  stat_function(
    fun = dnorm,
    args = list(mean = mean(stresids), sd = sd(stresids)),
    color = "red",
    linewidth = 1
  ) +
  labs(x = "Standardized Residuals",
       y = "Density",
       title = "Histogram of Standardized Residuals")

#Saving the Histogram
ggsave("Histogram_UV.png", width = 8, height = 6, dpi = 300)
```

```{r}
#Running KS Test
ks.test(stresids, "pnorm", mean = 0, sd = 1)
```

#### Equal Variance

```{r}
#Making Fitted Values Vs Standardized Residuals Plot
ggplot(data = data.frame(fit = fitted(uvlm), res = stresids),
       aes(x = fit, y = res)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "blue", size = 1.125) +
  labs(x = "Fitted",
       y = "Residuals", 
       title = "Fitted vs Residuals Plot") 

#Saving the Fitted Vs Residuals Plot
ggsave("Fitted_VS_Residuals_UV.png", width = 8, height = 6, dpi = 300)
```

```{r}
#Running BP Test
resid_df <- food |>
  mutate(r = stresids) |>
  select(-eating)
lmtest::bptest(formula = r ~ ., data = resid_df)
```
